---
title: "US Farmers' Markets Report"
author: "Daniel Pontoh"
date: "June 5, 2016"
output: html_document
code_folding: hide
runtime: shiny
---
*Data visuals in this reported were all made using data obtained from the [here](https://apps.ams.usda.gov/FarmersMarketsExport/ExcelExport.aspx) from the USDA.

```{r presetup, comment=FALSE, warning=FALSE, results='hide', message=FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(readxl)
library(reshape2)
library(stringi)

usfarmersmarkets <- read_csv("https://apps.ams.usda.gov/FarmersMarketsExport/ExcelExport.aspx")

uspopulation <- read_csv("https://www.census.gov/popest/data/national/totals/2015/files/NST-EST2015-alldata.csv")

#usfarmersmarkets <- read_csv("data/Export.csv")

#uspopulation <- read_csv("data/uspopulationdata.csv")
```





```{r datatransformation}

uspopulation <- select(uspopulation, NAME, CENSUS2010POP)

uspopulation$CENSUS2010POP <- as.numeric(uspopulation$CENSUS2010POP)

uspopulation <- uspopulation[6:57,]

virginislands <- c("Virgin Islands", as.numeric(106405))

uspopulation <- rbind(uspopulation, virginislands)

colnames(uspopulation) <- c("Var1", "Pop")

totusfm <- data.frame(table(usfarmersmarkets$State))

totusfm <- merge(uspopulation, totusfm, by = "Var1")


totusfm$lab <- as.character(totusfm$Freq)


totusfm$Pop <- as.numeric(totusfm$Pop)

totusfm$Freq <- as.numeric(totusfm$Freq)




totusfm$Pcp <- as.numeric(totusfm$Freq / totusfm$Pop) * 100000

totusfm$Pcp <- round(totusfm$Pcp, digits=1)
totusfm$pcplab <- as.character(totusfm$Pcp)
```


```{r plots}
gg <- ggplot(totusfm, aes(x=Var1, y=Freq, width = 0.75), color = "white")
gg <- gg + geom_bar(stat = "identity", fill="#fc8d62", color="#8da0cb")
gg <- gg + geom_text(aes(label = lab, hjust = -.25), size = 2.35)
gg <- gg + labs(x = "State/Territory",title= "Number of Farmers' Markets in Each State")
gg <- gg + theme(panel.grid.major.y = element_blank())
gg <- gg + theme(panel.grid.minor.x = element_blank())
gg <- gg + theme(panel.grid.minor.y = element_blank())
gg <- gg + theme(panel.border = element_blank())
gg <- gg + theme(panel.background = element_blank())
gg <- gg + theme(axis.ticks=element_blank())
gg <- gg + theme(plot.title = element_text(lineheight=.8, face="bold", size = 15, colour = "#66c2a5"))
gg <- gg + theme(axis.title.x = element_blank())
gg <- gg + theme(axis.title.y = element_text(face="bold", colour="#66c2a5", size=15))
gg <- gg + theme(axis.text.y  = element_text(size=8))
gg <- gg + theme(axis.text.x = element_blank())
gg <- gg + scale_y_continuous(expand = c(0,28.5))
gg <- gg + coord_flip()
gg

gg <- ggplot(totusfm, aes(x=Var1, y=Freq, width = 0.75), color = "white")
gg <- gg + geom_bar(stat = "identity", fill="#fc8d62", color="#8da0cb")
gg <- gg + geom_text(aes(label = lab, hjust = -.25), size = 2.35)
gg <- gg + labs(x = "State/Territory",title= "Number of Farmers' Markets in Each State")
gg <- gg + theme(panel.grid.major.y = element_blank())
gg <- gg + theme(panel.grid.minor.x = element_blank())
gg <- gg + theme(panel.grid.minor.y = element_blank())
gg <- gg + theme(panel.border = element_blank())
gg <- gg + theme(panel.background = element_blank())
gg <- gg + theme(axis.ticks=element_blank())
gg <- gg + theme(plot.title = element_text(lineheight=.8, face="bold", size = 15, colour = "#66c2a5"))
gg <- gg + theme(axis.title.x = element_blank())
gg <- gg + theme(axis.title.y = element_text(face="bold", colour="#66c2a5", size=15))
gg <- gg + theme(axis.text.y  = element_text(size=8))
gg <- gg + theme(axis.text.x = element_blank())
gg <- gg + scale_y_continuous(expand = c(0,28.5))
gg <- gg + scale_x_discrete(limits= totusfm$Var1[order(totusfm$Freq)])
gg <- gg + coord_flip()
gg




gg <- ggplot(totusfm, aes(x=Var1, y=Pcp, width = .75))
gg <- gg + geom_bar(stat = "identity", fill="#fc8d62", color="#8da0cb")
gg <- gg + geom_text(aes(label = pcplab, hjust = -.25), size = 2.35)
gg <- gg + labs(x = "State/Territory", title = "Number of Farmers' Markets per Capita")
gg <- gg + scale_y_continuous(expand = c(0,.85))
gg <- gg + theme(panel.grid.major.y = element_blank())
gg <- gg + theme(panel.grid.minor.x = element_blank())
gg <- gg + theme(panel.grid.minor.y = element_blank())
gg <- gg + theme(panel.border = element_blank())
gg <- gg + theme(panel.background = element_blank())
gg <- gg + theme(axis.ticks=element_blank())
gg <- gg + theme(plot.title = element_text(lineheight=.8, face="bold", size = 15, colour = "#66c2a5"))
gg <- gg + theme(axis.title.x = element_blank())
gg <- gg + theme(axis.title.y = element_text(face="bold", colour="#66c2a5", size=15))
gg <- gg + theme(axis.text.y  = element_text(size=8))
gg <- gg + theme(axis.text.x = element_blank())
gg <- gg + coord_flip()
gg



gg <- ggplot(totusfm, aes(x=Var1, y=Pcp, width = .75))
gg <- gg + geom_bar(stat = "identity", fill="#fc8d62", color="#8da0cb")
gg <- gg + scale_x_discrete(limits= totusfm$Var1[order(totusfm$Pcp)], expand = c(0,0))
gg <- gg + geom_text(aes(label = pcplab, hjust = -.25), size = 2.35)
gg <- gg + labs(x = "State/Territory", title = "Number of Farmers' Markets per Capita")
gg <- gg + scale_y_continuous(expand = c(0,.85))
gg <- gg + theme(panel.grid.major.y = element_blank())
gg <- gg + theme(panel.grid.minor.x = element_blank())
gg <- gg + theme(panel.grid.minor.y = element_blank())
gg <- gg + theme(panel.border = element_blank())
gg <- gg + theme(panel.background = element_blank())
gg <- gg + theme(axis.ticks=element_blank())
gg <- gg + theme(plot.title = element_text(lineheight=.8, face="bold", size = 15, colour = "#66c2a5"))
gg <- gg + theme(axis.title.x = element_blank())
gg <- gg + theme(axis.title.y = element_text(face="bold", colour="#66c2a5", size=15))
gg <- gg + theme(axis.text.y  = element_text(size=8))
gg <- gg + theme(axis.text.x = element_blank())
gg <- gg + coord_flip()
gg

summary(totusfm$Pcp)
```

```{r moredatatransformation}
usfm <- usfarmersmarkets[c(2, 9:11,21, 22,  24:58)]

usfm %>% 
  gather(var, value, Credit:WildHarvested) %>% 
  group_by(State, var) %>% 
  summarise(prop_Y = sum(value == "Y") / n()) %>% 
  spread(var, prop_Y) -> allstatespct

allstatespct$Average <- rowMeans(allstatespct[,-1])
allstatespct <- select(allstatespct, -Credit, -Organic, -SFMNP, - SNAP, -WIC, -WICcash)
allstatespct <- allstatespct[order(allstatespct$Average),]
```

```{r moregraphs, warning=FALSE}
gg <- ggplot(allstatespct, aes(x=State, y=Average, width = .75))
gg <- gg + geom_bar(stat = "identity", fill="#fc8d62", color="#8da0cb")
gg <- gg + labs(x = "State/Territory", title = "Average Farmers' Market Diversity")
gg <- gg + geom_text(aes(label = as.character(as.integer(allstatespct$Average*100)), hjust = -.25), size = 2.25)
gg <- gg + theme(panel.grid.major.y = element_blank())
gg <- gg + theme(panel.grid.minor.x = element_blank())
gg <- gg + theme(panel.grid.minor.y = element_blank())
gg <- gg + theme(panel.border = element_blank())
gg <- gg + theme(panel.background = element_blank())
gg <- gg + theme(axis.ticks=element_blank())
gg <- gg + theme(plot.title = element_text(lineheight=.8, face="bold", size = 15, colour = "#66c2a5"))
gg <- gg + theme(axis.title.x = element_blank())
gg <- gg + theme(axis.title.y = element_text(face="bold", colour="#66c2a5", size=15))
gg <- gg + theme(axis.text.y  = element_text(size=8))
gg <- gg + theme(axis.text.x = element_blank())
gg <- gg + coord_flip()
gg


gg <- ggplot(allstatespct, aes(x=State, y=Average, width = .75))
gg <- gg + geom_bar(stat = "identity", fill="#fc8d62", color="#8da0cb")
gg <- gg + labs(x = "State/Territory", title = "Average Farmers' Market Diversity")
gg <- gg + geom_text(aes(label = as.character(as.integer(allstatespct$Average*100)), hjust = -.25), size = 2.25)
gg <- gg + scale_x_discrete(limits= allstatespct$State[order(allstatespct$Average)], expand = c(0,0))
gg <- gg + theme(panel.grid.major.y = element_blank())
gg <- gg + theme(panel.grid.minor.x = element_blank())
gg <- gg + theme(panel.grid.minor.y = element_blank())
gg <- gg + theme(panel.border = element_blank())
gg <- gg + theme(panel.background = element_blank())
gg <- gg + theme(axis.ticks=element_blank())
gg <- gg + theme(plot.title = element_text(lineheight=.8, face="bold", size = 15, colour = "#66c2a5"))
gg <- gg + theme(axis.title.x = element_blank())
gg <- gg + theme(axis.title.y = element_text(face="bold", colour="#66c2a5", size=15))
gg <- gg + theme(axis.text.y  = element_text(size=8))
gg <- gg + theme(axis.text.x = element_blank())
gg <- gg + coord_flip()
gg

summary(allstatespct$Average)
```

```{r statediversityfunc}
statenames <- paste(levels(factor(usfarmersmarkets$State)))
selectInput("data", "", c(statenames))
```



```{r statediversitygraph}

renderPlot({
  d <- allstatespct[allstatespct$State == input$data,]
  d <- d[2,]
  d <- melt(d, id.vars="State")
  d <- d[2:3]
  d <- d[-30,]
  gg <- ggplot(d, aes(variable,value, fill = variable, width = .75))
  gg <- gg + geom_bar(stat = "identity", color="white")
  gg <- gg + labs(x = "Category", title = paste(input$data, "Farmers' Market Diversity"))
  gg <- gg + geom_text(aes(label = paste(as.character(as.integer(d$value*100)), "%"), hjust = -.30), size = 2.25)
  gg <- gg + scale_y_continuous(limits = c(0, 1), expand = c(.001,0))
  gg <- gg + theme(panel.grid.major.y = element_blank())
  gg <- gg + theme(panel.grid.minor.x = element_blank())
  gg <- gg + theme(panel.grid.minor.y = element_blank())
  gg <- gg + theme(panel.border = element_blank())
  gg <- gg + theme(panel.background = element_blank())
  gg <- gg + theme(axis.ticks=element_blank())
  gg <- gg + theme(plot.title = element_text(lineheight=.8, face="bold", size = 15, colour = "#66c2a5"))
  gg <- gg + theme(axis.title.x = element_blank())
  gg <- gg + theme(axis.title.y = element_text(face="bold", colour="#66c2a5", size=18))
  gg <- gg + theme(axis.text.y  = element_text(size=8))
  gg <- gg + theme(axis.text.x = element_blank())
  gg <- gg + theme(plot.title = element_text(size = 16))
  gg <- gg + theme(legend.position = "none")
  gg <- gg + coord_flip()
  gg
}, res = 90)
```

```{r categorydiversityfunc}
selectInput("data2", "", choices = c("Bakedgoods" = 2,
  "Beans" = 3,
  "Cheese" = 4,
  "Coffee" = 5,
  "Crafts" = 6,
  "Eggs" = 7,
  "Flowers" = 8,
  "Fruits" = 9,
  "Grains" = 10,
  "Herbs" = 11,
  "Honey" = 12,
  "Jams" = 13,
  "Juices" = 14,
  "Maple" = 15,
  "Meat" = 16,
  "Mushrooms" = 17,
  "Nursery" = 18,
  "Nuts" = 19,
  "PetFood" = 20,
  "Plants" = 21,
  "Poultry" = 22,
  "Prepared" = 23,
  "Seafood" = 24,
  "Soap" = 25,
  "Tofu" = 26,
  "Trees" = 27,
  "Vegetables"= 28,
  "WildHarvested" = 29,
  "Wine" = 30))

```

```{r categorydiversitygraph}

renderPlot({
  b <- c(as.numeric(1),as.numeric(input$data2))
  a <- allstatespct[,b]
  colnames(a) <- c("state", "variable")
gg <- ggplot(a, aes(state,variable))
gg <- gg + geom_bar(stat = "identity", aes(fill = state), color = "white")
gg <- gg + labs(x = "State/Territory", title = "Farmers' Market Diversity")
  gg <- gg + geom_text(aes(label = paste(as.character(as.integer(a$variable*100)), "%"), hjust = -.30), size = 2.25)
  gg <- gg + theme(panel.grid.major.y = element_blank())
  gg <- gg + theme(panel.grid.minor.x = element_blank())
  gg <- gg + theme(panel.grid.minor.y = element_blank())
  gg <- gg + theme(panel.border = element_blank())
  gg <- gg + theme(panel.background = element_blank())
  gg <- gg + theme(axis.ticks=element_blank())
  gg <- gg + theme(plot.title = element_text(lineheight=.8, face="bold", size = 15, colour = "#66c2a5"))
  gg <- gg + theme(axis.title.x = element_blank())
  gg <- gg + theme(axis.title.y = element_text(face="bold", colour="#66c2a5", size=18))
  gg <- gg + theme(axis.text.y  = element_text(size=8))
  gg <- gg + theme(axis.text.x = element_blank())
  gg <- gg + theme(plot.title = element_text(size = 16))
  gg <- gg + theme(legend.position = "none")
  gg <- gg + coord_flip()
  gg
}, red = 90)